---
output:
  bookdown::github_document2:
    fig_caption: true
    toc: true
bibliography: https://api.citedrive.com/bib/1306e547-b1d9-497f-a7c1-120c6d3d6ad7/references.bib?x=eyJpZCI6ICIxMzA2ZTU0Ny1iMWQ5LTQ5N2YtYTdjMS0xMjBjNmQzZDZhZDciLCAidXNlciI6ICI4NjQ0IiwgInNpZ25hdHVyZSI6ICJmMjFmNGMwNjZhOWYxN2I0YmRlZTdjNWU2ZmY1N2Y5OTA1MGRjNTlmZGIwN2ZlMzFiNTc2YzY0ZjI1MzBlYzA3In0=/bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  error = FALSE
)
```

# Pandemic_risk_and_deforestation

<!-- badges: start -->
<!-- badges: end -->

The goal of the *Pandemic_risk_and_deforestation* is to document the code and datasets used for the mansucript *Pandemic risk and deforestation policies in the world tropics* this repo goes section by section with the code

## load necessary packages:

first we load the necessary packages

```{r loadpackages}
library(terra)
library(dplyr)
library(broom)
library(geodata)
library(purrr)
library(tidyterra)
library(ggplot2)
```


## Mammal species richness, abundance and number of infected individuals 

We used IUCN range maps for mammalian species in the Orders Chiroptera, Primate, Rodentia and Ungulata gathered from the IUCN red list (https://www.iucnredlist.org/resources/spatial-data-download) in raster format to calculate species richness ($S$) for each order within each cell. To transform number of species into number of individuals, we used models fitted to empirical data obtained from a world database of species richness and abundance in local communities throughout the world (The Ecological Register [@alroy2015shape].

```{r AlroyData}
Data <- readRDS("Alroy_Data.rds") |>  
    dplyr::filter(Goods_U > 0.9, count > 100) |>  
    dplyr::mutate(LogS = log(Richness), LogA = log(Fishers_Alpha))
```

From these relationships (for values of Goods_U > 0.9 and counts of more of a 100), we then calculated the expected abundance for each cell $S$

```{r NLSModel}
powerlaw.model <- nls(Fishers_Alpha~a*Richness^y, start= list(y=0, a = 1), data =Data)
```

see results in table \@ref(tab:tableNLS)

```{r tableNLS, echo = FALSE}
knitr::kable(broom::tidy(powerlaw.model), caption = "Parameters from the non linear model for the power law of fishers alpha", digits = 3)
```

## Population density

We calculated the population density based on [@ciesin2018], at 2.5 min resolution, and then multiply them by the area to get total population, that is then transformed to equal area projection to make it consistent

```{r density, cache=TRUE, include=FALSE}
population <- seq(2000, 2020, by = 5) |>
  purrr::map(~geodata::population(year = .x, res = 2.5, path = getwd())) |> 
  purrr::reduce(c)

# Multiply by area

population <- population*terra::cellSize(population[[1]], unit = "km")

# Transform to equal area

population <- terra::project(population, "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")

names(population) <- paste0("Year", seq(2000, 2020, by = 5))

# save to cloud optimized geotif
terra::writeRaster(x = population, filename = "Population.tif", overwrite = TRUE, gdal = c("COMPRESS=DEFLATE", "TFW=YES", "of=COG"))
```

```{r densityshow, eval=FALSE}
population <- seq(2000, 2020, by = 5) |>
  purrr::map(~geodata::population(year = .x, res = 2.5, path = getwd())) |> 
  purrr::reduce(c)

# Multiply by area

population <- population*terra::cellSize(population[[1]], unit = "km")

# Transform to equal area

population <- terra::project(population, "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")

names(population) <- paste0("Year", seq(2000, 2020, by = 5))

# save to cloud optimized geotif
terra::writeRaster(x = population, filename = "Population.tif", overwrite = TRUE, gdal = c("COMPRESS=DEFLATE", "TFW=YES", "of=COG"))
```

This looks like this in the log scale in figure \@ref(fig:ShowPopulation)

```{r ShowPopulation, cache=TRUE, fig.cap= "Population by pixel in the world", echo=FALSE}
population <- terra::rast("Population.tif")

ggplot() + 
  geom_spatraster(data = population) + 
  facet_wrap(~lyr) +
  scale_fill_wiki_c(trans = "log", labels = scales::comma) +
  theme_bw() 
```


# Used Packages and Versions

```{r echo=FALSE}
# Load knitr package
library(knitr)

# Get loaded packages information
loaded_packages <- sessionInfo()$otherPkgs

# Extract package names and versions
package_info <- data.frame(
  Package = names(loaded_packages),
  Version = sapply(loaded_packages, function(x) x$Version),
  stringsAsFactors = FALSE
)

# Remove base and recommended packages
package_info <- package_info[!(package_info$Package %in% c("base", "recommended")), ]
```


```{r list, results='asis', echo = F}
# Print package information in Markdown format
cat(paste0("* **", package_info$Package, "**: ", package_info$Version, "\n"))
```


In this code, `sessionInfo()$otherPkgs` is used to extract the loaded packages during the rendering of the R Markdown document. We then extract the package names and versions from this information and format it into a Markdown list, excluding base and recommended packages.


# References